Fixing Vulnerabilities Potentially Hinders Maintainability

### 2021-01-26 response/response.tex

\end{revcomment}

\Us Thank you for suggesting Marco's paper. Our paper uses a model 
for measuring maintainability that was published previously by one of the co-authors
of this paper [Cruz2019]. The entire methodology and calculations 
were actually validated by Marco.

The model created by Marco focus on measuring maintainability of fine-grained 
changes using 5 of the same guidelines we already use. The SIG-MM model in 
Marco's paper is the model behind Better Code Hub.

We mention Marco's work in our related work:
\begin{quote}
    Recent work,
proposed a new maintainability model to 
measure fine-grained code changes by adapting/extending the BCH model[diBiase2019].
The present work uses the same model (SIG-MM), considers more guidelines and focuses 
solely on evaluating the impact of security patches on software maintainability.
\end{quote}

@@

[Alves2010] T. L. Alves and C. Ypma and J. Visser. Deriving metric thresholds from 
benchmark data. IEEE International Conference on Software Maintenance in 2010.

[Alves2011] T. L. Alves and J. P. Correia and J. Visser. Benchmark-Based Aggregation 
of Metrics to Ratings. Joint Conference of the 21st International Workshop on Software 
Measurement and the 6th International Conference on Software Process and Product Measurement
in 2011.

[Baggen2012] R. Baggen, J. P. Correia, K. Schill, J. Visser. Standardized code quality 
benchmarking for improving software maintainability. Software Quality Journal in 2012.

\begin{revcomment}[2.8]
    Section 3\\
    \begin{itemize}
        \item Page 7, line 4 - I don't think "resemble" is the word you are looking for - but I don't 
        know what you are trying to say so I can't make a suggestion.
    \end{itemize}
\end{revcomment}

\Us Thank you for raising this issue. In the sentence ``The codebases that resemble to the commits of our datasets 
are collected before the analysis performed by BCH.'', we mean that the codebases of each commit in 
our dataset were analyzed by BCH. We improved the text for:

\begin{quote}
    BCH evaluates the codebase available in the default branch of a GitHub project. 
  We created a tool that pulls the codebase of each commit of our dataset 
  to a new branch; it sets the new branch as the default branch; it runs 
  the BCH analysis in the codebase; and, finally saves the BCH metrics
  results.
\end{quote}

\begin{revcomment}[2.9]
    Section 3\\
    \begin{itemize}
        \item Please explain your repeatable methodology to "clean the dataset and select the most relevant and 
        compatible projects" (page 8, line 14) that brings the patches from 1282 to 969.  From 969 to 866 patches - 
        you just could not classify?  Please explain.
    \end{itemize}
\end{revcomment}

\Us As explained in comment $2.6$, the $1282$ commits refer only to the Ponta et al. dataset (a total of $624$ 
patches). In total, we analyzed $1300$ patches. However, BCH as some limitations such as
analyzing projects with very large codebases and was uncapable to analyze some patches. In addition,
we also detected floss-refactorings which we decided not to consider. All of this is explained
in more detail later in Section 3.4.

\begin{quote}
    Security patches can be performed through one commit, 
several consecutive commits, or commit(s) interleaved with more 
programming activities (floss-refactoring). This study considers 
mainly single-commit patches. Yet, a small percentage of data points 
had more than one commit involved in the patch. We manually 
inspected these cases and $23$ floss-refactorings were identified 
and disregarded since it would not be fair to measure the 
maintainability of these cases where other programming activities 
are involved.

For projects with large codebases, the results retrieved for 
\emph{Keep Your Codebase Small} were not viable because they 
retrieved values above the limit set by BCH ($20$ Person-years). The 
\emph{Automated Tests} guideline was also not considered since the 
tool does not contemplate two of the most important techniques to 
security testing: vulnerability scanning and penetration testing. 
Instead, it only contemplates unit testing. Due to BCH limitations, 
we detected a few data points that retrieved incorrect overall 
calculations. Those data points were disregarded from the study.
\end{quote}

Those are the cases that were tossed from the initial $1300$ patches.

Regarding the second question, thank you for pointing out ... (still working on this)

\begin{revcomment}[2.10]
    Section 3\\
    \begin{itemize}
        \item Page 8, line 33 "regular commit" - pick baseline or regular (or non-security) commits and use it always.  
        I don't understand the sentence "The baseline dataset is generated from the security commits dataset."   
        Do you mean you extract these commits from the projects in the security commits dataset?
    \end{itemize}
\end{revcomment}

\Us Yes, for each security commit we collected a random regular change from the same project.

@@

\begin{revcomment}[2.11]
    Section 3\\
    \begin{itemize}
        \item Section 3.2 is confusing enough that I'm not sure of the analysis - 
        I will need to review the revision that you would submit to gain confidence.  
        Are you saying you want your non-security commits (regular commits) to be about 
        the same size as a given security commit so you are using this criteria to choose 
        your regular commits?  Justify this reasoning and make the methodology clear.  
        I would have expected that you randomly collected 1300 non-security commits to compare 
        with the 1300 security commits. Why constrain to make them be of "similar" size?  
        Why not allow the characteristics of regular versus security commits to be what 
        they are without such curating?
    \end{itemize}
\end{revcomment}

\Us Yes, we are looking for regular changes with the same size as security commits. We 
have collected a baseline of random changes only (no size restrictions) in the past and 
performed the same evaluation. We now present the results for both baselines: \textit{size-baseline}, where we consider
size appproximation; and, \textit{random-baseline}, where we consider all the regular
changes characteristics.

@@

\begin{revcomment}[2.14]
    Section 3\\
    \begin{itemize}
        \item Page 11, Line 42 - I think you are saying you removed these two guidelines 
        from the analysis of all projects (as shown in Figure 4) but this paragraph seems 
        to say this was only for projects with large code bases.
    \end{itemize}
\end{revcomment}

\Us working on this

\begin{revcomment}[2.15]
    Section 3\\
    \begin{itemize}
        \item Page 11  line 48.  How many were disregarded?
    \end{itemize}
\end{revcomment}

\Us Due to BCH limitations, $308$ data points were disregarded.

@@

\begin{revcomment}[3.1]

    The paper addresses a relevant and up to date topic. 
    The authors have analyzed a large set of project wrt to 
    quality assessment focused on vulnerabilities in line with 
    ISO25000 standard. There are some points worth considering: 
    - authors should better motivate the choice for BCH as web 
    based source code analysis service with respect to other 
    options that are available in literature. Tools such as Kiuwan, 
    sonar cloud, codify etc are also able to provide specific metrics 
    on maintainability and security, vulnerability features of code. 
    Furthermore Kiuwan is also an example designed based on the ISO25000 
    standard. Authors discuss the need for tools for risk assessment in 
    the study implications (section5) but fail to assess why their choice 
    falls on BCH. Furthermore all tools provide different conceptualization 
    and evaluation criteria for calculating quality characteristics such as Vulnerabilities 
    and Security. Authors should also explicit how these quality characteristics are measured in BCH.

\end{revcomment}

\Us We kindly thank you the reviewer for the comments. We added a sub Section to 
the paper called \textit{Why Better Code Hub?}. We compare BCH with other tools
that perform code analysis and motivate why we find BCH to be the best fit for us.

### 2021-01-26 paper/emse20.tex
more fixes

As ISO does not provide any specific guidelines/formulas to 
calculate maintainability, we resort to Software Improvement Group 
(SIG\footnote{SIG's website: 
\url{https://www.sig.eu/} (Accessed on \today{})})'s web-based source 
code analysis service Better Code Hub (BCH)\footnote{BCH's 
website: \url{https://bettercodehub.com/} (Accessed on \today{})}  
to compute the software compliance with a set of $10$ 
guidelines/metrics to produce quality software based in ISO/IEC 
$25010$~\cite{Visser:2016:OREILLY}. SIG 
has been helping business and technology leaders drive their organizational 
objectives by fundamentally improving the health and security of 
their software applications for more than 20 years. Their 
models are scientifically proven and certified~\cite{5609747,6113040,baggen2012}.

@@

In this section, we discuss the methodology used to measure
the impact of security patches on the maintainability of open-source
software. The methodology comprises the following steps, as 
illustrated in Figure~\ref{fig:met}.
%
\begin{enumerate}
	\item Combine the datasets from related work that classifies
	the activities of developers addressing security-oriented 
	patches~\cite{Reis:2017:IJSSE,10.1109/MSR.2019.00064}.
%
\item After combining both datasets, we removed the duplicates
and filtered out every floss-refactoring.
%
	\item
	Extract relevant data (e.g., owner and name of
	the repository, sha key of the vulnerable version, sha key of 	
	the fixed version and codebases) from the combined dataset 	
	containing $1300$ security patches collected from open-source 	
	software available on GitHub.
%
  \item A baseline of regular commits was randomly collected (i.e., for 
  each security commit a random commit from the same project was selected) from 
	the list of projects of the final dataset to evaluate the impact 
	of regular commits on the maintainability of open-source 
	software.
%
  \item Use the Software Improvement Group's\footnote{SIG's website: 
  \url{https://www.sig.eu/} (Accessed on \today{})} web-based source 
  code analysis service \emph{Better Code Hub} (BCH)\footnote{BCH's 
  website: \url{https://bettercodehub.com/} (Accessed on \today{})} 
  to quantify maintainability for both security and regular commits. 
  BCH evaluates the codebase available in the default branch of a GitHub project. 
  We created a tool that pulls the codebase of each commit of our dataset 
  to a new branch; it sets the new branch as the default branch; it runs 
  the BCH analysis in the codebase; and, finally saves the BCH metrics
  results.
\end{enumerate}
%
\begin{figure}[h]
	\centering 	
    \begin{adjustwidth}{-0.5cm}{-0.5cm}  
      \vspace{-2.5cm} 
	\includegraphics[width=1.1\textwidth]{figures/methodology.pdf}
  \vspace{-3cm} 
  \caption{Study Methodology}
	\label{fig:met}
	 \end{adjustwidth}
\end{figure}
%
\subsection{Datasets}
%

We use a combined dataset of $1300$ security patches which is 
the outcome of mining and manually inspecting a total of $312$ 
GitHub projects. The combined dataset integrates two different
works: SecBench~\cite{Reis:2017:IJSSE} and Ponta et al.~\cite{10.1109/MSR.2019.00064}.

Reis and Abreu 
($2017$) mined open-source software aiming at the extraction of 
real---created by developers---patches of security vulnerabilities 
to test and assess the performance of static analysis 
tools~\cite{Reis:2017:IJSSE} since using hand-seeded test cases or 
mutations can lead to misleading assessments of the capabilities of 
the tools~\cite{just2014mutants}. The study yielded a dataset of 
$676$ patches for $16$ different security vulnerability types, dubbed as SecBench. 
The vulnerability types are based on the OWASP Top $10$ of 
$2013$~\cite{oswap:2013} and OWASP Top $10$ of 
$2017$~\cite{oswap:2017}. Each test case of the dataset is a 
triplet: the commit before the patching (\emph{sha-p}), the commit responsible
for the patching (\emph{sha}), and the snippets of code that differ from one 
version to another (typically, called \emph{diffs})---where one 
can easily review the code used to fix the vulnerability. 
%

Ponta et al. ($2019$) tracked the \url{pivotal.io} website for 
vulnerabilities from $2014$ to $2019$. For each new vulnerability, 
the authors manually searched for references to commits involved in 
the patch on the National Vulnerability Database (NVD) website. 
However, $70\%$ of the vulnerabilities did not have any references 
to commits. Thus, the authors used their expertise to locate the 
commits in the repositories. This technique yielded a dataset of 
$624$ patches~\cite{10.1109/MSR.2019.00064} and $1282$ commits. 
One patch can have multiple commits 
assigned. To fit the dataset in our methodology, we located the 
first and last commits used to patch the vulnerability. For these 
cases, we used the GitHub API to retrieve automatically the dates of 
the commits. Then, for each patch, the group of commits was ordered 
from the oldest commit to the newest one. We assumed the last commit 
(newest one) as the fix (\emph{sha}) and the parent of the 
first commit (oldest commit) as the vulnerable version 
(\emph{sha-p}). 
%

In this study, we focus on computing the maintainability of the 
commits before and after the security patching to evaluate if its 
impact was positive, negative or none. The $1300$ patches in the 
dataset were analyzed using the BCH toolset to calculate their 
maintainability reports. Due to the limitations of BCH, in particular, 
lack of language support and project size, we had
to clean the dataset and select the most relevant and compatible projects
with BCH. The final dataset used in this 
paper comprises $969$ security patches from $260$ projects. We used the 
Common Weakness Enumeration (CWE) taxonomy to classify each vulnerability. 
For instance,  the \texttt{Fix CVE-2014-1608: mc\_issue\_attachment\_get SQL injection} 
(\texttt{00b4c17}\footnote{CVE-$2014$-$1608$ details available at 
\url{https://github.com/mantisbt/mantisbt/commit/00b4c17088fa56594d85fe46b6c6057bb3421102} 
(Accessed on \today)}) is a \emph{CWE-89: Improper Neutralization of Special Elements used 
in an SQL Command ('SQL Injection')}\footnote{CWE-$89$ details available at \url{https://cwe.mitre.org/data/definitions/89.html} 
(Accessed on \today)} according to the CWE taxonomy.
We were able to classify a total of $866$ patches using the CWE taxonomy
where the CWEs for $536$ patches were automatically scraped from the National 
Vulnerability Dataset (NVD); while for the other $330$ patches, the CWEs were manually 
chosen by the authors following the \emph{Research Concepts} CWE's list.

@@

%     \item \textbf{CWE-89: Improper Neutralization of Special Elements used in an SQL Command/SQL Injection (23 cases).} When developers do not keep untrusted data
%     separate from SQL queries. If an attacker sends a command that
%     exploits the syntax of the SQL interpreter, then SQL injection attack is possible.
% %
%     \item \textbf{CWE-611: Improper Restriction of XML External Entity Reference (24 cases).}
% 	XML documents are exchanged through the web containing entities with URIs
% 	that resolve to local/external files. Thus, when XML parsers are not well configured,
% 	attackers have allowed to directly those files.
% %
%     \item \textbf{CWE-20: Improper Input Validation (76 cases).} When developers
% 	forget to validate the inputs of an application, attackers may have control
% 	of the control/data flow of the program.
% %
%     \item \textbf{CWE-200: Information Exposure (39 cases).} Application's logs are one way of intentional/unintentional disclosure information
% 	to attackers. Many times attackers get access to logs when sould not be authorized to.
% %
% 	\item \textbf{CWE-352: Cross-Site Request Forgery/CSRF (22 cases).} Poor session tokens generation
%     and management usually allow attackers to send forged HTTP requests
%     including authentication information from the victim to the vulnerable web
%     application.
% %
% 	\item \textbf{CWE-264: Permissions, Privileges, and Access Controls (59 cases).} Incorrectly
% 	    implemented functionalities related to authentication and session
% 	    management, allowing an attacker to gain access to session tokens,
% 	    passwords, keys, and other sensitive data.
% %
% 	\item \textbf{CWE-399: Resource Management Errors (22 cases).} Resource management issues are found more frequently
% 		in programming languages that do not manage memory automatically (e.g., C/C++
% 	    and Objective-C), i.e., where developers are responsible for
% 	    handling it instead. It is one of the main causes of DoS attacks.
% %
% 	\item \textbf{CWE-79: Improper Neutralization of Input During Web Page Generation/Cross-site Scripting (62 cases)} Lack of proper validation or escaping
% 	    allows attackers to submit untrusted data to web browsers through malicious
% 	    scripts that can hijack the user sessions or redirect the user to malicious
% 	    sites.
% %
% 	\item \textbf{CWE-22: Improper Limitation of a Pathname to a Restricted Directory/Path Traversal (37 cases).}  When developers forget to neutralize special elements within the pathnames to files/directories, attackers
% 	may leverage this flaw to resolve to a location outside of the restricted directory and gain access to documents
% 	or entire repositories.
% %
% 	\item \textbf{Miscellaneous (673 cases)} This pattern comprises
% 		several other security patches that do not have a CWE assigned and patterns that
% 		do not satisfy the Wilcoxon test size requirement of more than $20$ patches.
\subsection{Security vs. Baseline Commits}
Previous studies attempted to measure the impact of regular commits 
on open-source software maintainability~\cite{HEGEDUS2018313}. 
However, there is no previous work focused on comparing the impact 
of security patches with regular commits on maintainability.
We analyze the maintainability of regular commits (i.e., commits not 
related to security patches) and use them as a baseline.
The baseline dataset is generated from the security commits dataset. 
As for the security patches, for each regular fix or refatoring, we also need the commit performing the regular change (\emph{sha-reg}) and version of 
the software before the change (\emph{sha-reg-p}). For each security patch, 
a \emph{diff} with an approximate size was fetched to a regular 
change. Due to the complexity of some patches, it was not possible 
to find patches with the exact number of added and deleted lines. 
Our search uses the security patch \emph{diff} to look for a 
\emph{diff} of the same size. The algorithm integrates the following 
steps:

@@

The pair of the random commit and its parent is accepted if the 
\emph{diff} size fits in the range size. This range widens every 
$10$ attempts to search for a change with the same size. We 
originate the regular commits from the security commits to ensure 
that differences in maintainability are not a consequence of 
characteristics of different projects.

@@

\subsection{Maintainability Analysis}

In this research, we follow a very similar methodology to the one 
presented in previous work on the maintainability of energy-oriented 
fixes~\cite{8919169}. As said before, the web-based source code 
analysis service \emph{Better Code Hub} (BCH) is used to collect the 
maintainability reports of the patches of each project. 
Table~\ref{tab:guidelines} presents the $10$ guidelines proposed
by BCH's authors for delivering software that is not difficult to
maintain~\cite{Visser:2016:OREILLY} and, maps each guideline to the 
metric calculated by BCH. Theses guidelines are calculated using the 
metrics presented in~\cite{criteria:2017} and are also briefly 
explained on Table~\ref{tab:guidelines}. 
During each guideline evaluation, the tool determines the compliance 
towards one guideline by establishing limits for the percentage of 
code allowed to be in each of the $4$ risk severity levels
(\emph{low risk}, \emph{medium risk}, \emph{high risk}, and 
\emph{very high risk}). If the project does not violate those 
thresholds, then the BCH considers that the code is compliant with 
a guideline. These thresholds are determined by BCH using their own
data/experience---using open-source and closed software systems. If 
a project is compliant with a guideline, it means that it is at 
least $65\%$ better than the software used by BCH to calculate the 
thresholds\footnote{Check the answer to \emph{How can I adjust the 
threshold for passing/not passing a guideline?} at
\url{https://bettercodehub.com/docs/faq} (Accessed on \today{})}.

\begin{table}[h]
\centering
\scriptsize
	\caption{Guidelines to produce maintainable code.}
\begin{tabular}{L{2cm}L{4.5cm}L{4cm}}

@@

For projects with large codebases, the results retrieved for 
\emph{Keep Your Codebase Small} were not viable because they 
retrieved values above the limit set by BCH ($20$ Person-years). The 
\emph{Automated Tests} guideline was also not considered since the 
tool does not contemplate two of the most important techniques to 
security testing: vulnerability scanning and penetration testing. 
Instead, it only contemplates unit testing. Due to BCH limitations, 
we detected a few data points that retrieved incorrect overall 
calculations. Those data points were disregarded from the study.

@@

Many studies have investigated the relationship between patches and
software quality. Previous work focused on object-oriented metrics has evaluated the
impact of patches and exhibited proof that quantifying effectively the
impact of patches on maintainability may help to choose the appropriate
patch type~\cite{1167822}. In contrast to this work, Hegedus et
al.~\cite{HEGEDUS2018313} did not select particular metrics to assess the effect
of patches. Instead, statistical tests were used to find the metrics that
have the potential to change significantly after patches. Studying the evolution 
of maintainability issues during the development of Android apps, Malavolta et al. ($2018$)~\cite{8530041}
discovered that maintainability decreases over time. Palomba et al.
($2018$)~\cite{Palomba:2018:DIM:3231288.3231337} exhibits proof that code smells
should be carefully monitored by programmers since there is a high correlation
between maintainability aspects and proneness to changes/faults. In 2019, Cruz et 
al.~\cite{8919169} proposed a formula to calculate maintainability 
based on the BCH's guidelines and measured the impact of energy-oriented fixes 
on software maintainability. Recent work,
proposed a new maintainability model to 
measure fine-grained code changes by adapting/extending the BCH model~\cite{8785997}.
The present work uses the same model (BCH) and focuses 
solely on evaluating the impact of security patches on software maintainability.

>>>>

As ISO does not provide any specific guidelines/formulas to 
calculate maintainability, we resort to Software Improvement Group 
(SIG\footnote{SIG's website: 
\url{https://www.sig.eu/} (Accessed on \today{})})'s web-based source 
code analysis service Better Code Hub (BCH)\footnote{BCH's 
website: \url{https://bettercodehub.com/} (Accessed on \today{})}  
to compute the software compliance with a set of $10$ 
guidelines/metrics to produce quality software based in ISO/IEC 
$25010$~\cite{Visser:2016:OREILLY}. SIG 
has been helping business and technology leaders drive their organizational 
objectives by fundamentally improving the health and security of 
their software applications for more than 20 years. Their 
models are scientifically proven and certified~\cite{4335232,5609747,6113040,baggen2012}.

@@

In this section, we discuss the methodology used to measure
the impact of security patches on the maintainability of open-source
software. The methodology comprises the following steps, as 
illustrated in Figure~\ref{fig:met}.
%
\begin{enumerate}
	\item Combine the datasets from related work that classify
	the activities of developers addressing security-oriented 
  patches~\cite{reis2017secbench,10.1109/MSR.2019.00064}. The
  duplicated patches were tossed (only $2$ were found).
%
	\item
	Extract relevant data (e.g., owner and name of
	the repository, sha key of the vulnerable version, sha key of 	
	the fixed version) from the combined dataset 	
	containing $1300$ security patches collected from open-source 	
	software available on GitHub.
%
  \item Two baselines of regular changes were collected:
  \textit{random-baseline} (for each security commit, a random change 
  was collected from the same project) and \textit{size-baseline} (for 
  each security commit, a random change with the same size
  was collected from the same project). Our goal is to evaluate the impact 
	of regular changes on the maintainability of open-source 
	software.
%
  \item Use the Software Improvement Group (SIG)'s web-based source 
  code analysis service \emph{Better Code Hub} (BCH)
  to quantify maintainability for both security and regular commits. 
  BCH evaluates the codebase available in the default branch of a GitHub project. 
  We created a tool that pulls the codebase of each commit of our dataset 
  to a new branch; it sets the new branch as the default branch; it runs 
  the BCH analysis in the codebase; and, finally saves the BCH metrics
  results.
\end{enumerate}
%
\begin{figure}[h]
	\centering 	
    \begin{adjustwidth}{-0.5cm}{-0.5cm}  
      \vspace{-2.5cm} 
	\includegraphics[width=1.1\textwidth]{figures/methodology.pdf}
  \vspace{-3cm} 
  \caption{Study Methodology}
	\label{fig:met}
	 \end{adjustwidth}
\end{figure}
%
\subsection{Datasets}
%

We use a combined dataset of $1300$ security patches which is 
the outcome of mining and manually inspecting a total of $312$ 
GitHub projects. The combined dataset integrates two different
works: SECBENCH~\cite{reis2017secbench,Reis:2017:IJSSE} and Ponta et al.~\cite{10.1109/MSR.2019.00064}.

Reis and Abreu 
($2017$) mined open-source software aiming at the extraction of 
real---created by developers---patches of security vulnerabilities 
to test and assess the performance of static analysis 
tools~\cite{reis2017secbench,Reis:2017:IJSSE} since using hand-seeded test cases or 
mutations can lead to misleading assessments of the capabilities of 
the tools~\cite{just2014mutants}. The study yielded a dataset of 
$676$ patches for $16$ different security vulnerability types, dubbed as SECBENCH. 
The vulnerability types are based on the OWASP Top $10$ of 
$2013$~\cite{oswap:2013} and OWASP Top $10$ of 
$2017$~\cite{oswap:2017}. Each test case of the dataset is a 
triplet: the commit before the patching (\emph{sha-p}), the commit responsible
for the patching (\emph{sha}), and the snippets of code that differ from one 
version to another (typically, called \emph{diffs})---where one 
can easily review the code used to fix the vulnerability. 
%
Ponta et al. ($2019$) tracked the \url{pivotal.io} website for 
vulnerabilities from $2014$ to $2019$. For each new vulnerability, 
the authors manually searched for references to commits involved in 
the patch on the National Vulnerability Database (NVD) website. 
However, $70\%$ of the vulnerabilities did not have any references 
to commits. Thus, the authors used their expertise to locate the 
commits in the repositories. This technique yielded a dataset of 
$624$ patches~\cite{10.1109/MSR.2019.00064} and $1282$ commits---one 
patch can have multiple commits assigned. To fit the dataset in our methodology, we located the 
first and last commits used to patch the vulnerability. For these 
cases, we used the GitHub API to retrieve automatically the dates of 
the commits. Then, for each patch, the group of commits was ordered 
from the oldest commit to the newest one. We assumed the last commit 
(newest one) as the fix (\emph{sha}) and the parent of the 
first commit (oldest commit) as the vulnerable version 
(\emph{sha-p}). 
%

In this study, we focus on computing the maintainability of the 
commits before and after the security patching to evaluate if its 
impact was positive, negative or none. The $1300$ patches in the 
dataset were analyzed using the BCH toolset to calculate their 
maintainability reports. Due to the limitations of BCH (in particular, 
lack of language support and project size) and the presence of
floss-refactorings, $331$ patches were tossed---explained in 
more detail in Section~\ref{sec:main_analysis}.
The final dataset used in this 
paper comprises $969$ security patches from $260$ projects. We used the 
Common Weakness Enumeration (CWE) taxonomy to classify each vulnerability. 
For instance,  the \texttt{Fix CVE-2014-1608: mc\_issue\_attachment\_get SQL injection} 
(\texttt{00b4c17}\footnote{CVE-$2014$-$1608$ details available at 
\url{https://github.com/mantisbt/mantisbt/commit/00b4c17088fa56594d85fe46b6c6057bb3421102} 
(Accessed on \today)}) is a \emph{CWE-89: Improper Neutralization of Special Elements used 
in an SQL Command ('SQL Injection')}\footnote{CWE-$89$ details available at \url{https://cwe.mitre.org/data/definitions/89.html} 
(Accessed on \today)} according to the CWE taxonomy.
We were able to classify a total of $866$ patches \Sofia{not sure why i did not classified all of them}using the CWE taxonomy
where the CWEs for $536$ patches were automatically scraped from the National 
Vulnerability Dataset (NVD); while for the other $330$ patches, the CWEs were manually 
chosen by the authors following the \emph{Research Concepts} CWE's list.

@@

%
% \begin{itemize}
%     \item \textbf{CWE-89: Improper Neutralization of Special Elements used in an SQL Command/SQL Injection (23 cases).} When developers do not keep untrusted data
%     separate from SQL queries. If an attacker sends a command that
%     exploits the syntax of the SQL interpreter, then SQL injection attack is possible.
% %
%     \item \textbf{CWE-611: Improper Restriction of XML External Entity Reference (24 cases).}
% 	XML documents are exchanged through the web containing entities with URIs
% 	that resolve to local/external files. Thus, when XML parsers are not well configured,
% 	attackers have allowed to directly those files.
% %
%     \item \textbf{CWE-20: Improper Input Validation (76 cases).} When developers
% 	forget to validate the inputs of an application, attackers may have control
% 	of the control/data flow of the program.
% %
%     \item \textbf{CWE-200: Information Exposure (39 cases).} Application's logs are one way of intentional/unintentional disclosure information
% 	to attackers. Many times attackers get access to logs when sould not be authorized to.
% %
% 	\item \textbf{CWE-352: Cross-Site Request Forgery/CSRF (22 cases).} Poor session tokens generation
%     and management usually allow attackers to send forged HTTP requests
%     including authentication information from the victim to the vulnerable web
%     application.
% %
% 	\item \textbf{CWE-264: Permissions, Privileges, and Access Controls (59 cases).} Incorrectly
% 	    implemented functionalities related to authentication and session
% 	    management, allowing an attacker to gain access to session tokens,
% 	    passwords, keys, and other sensitive data.
% %
% 	\item \textbf{CWE-399: Resource Management Errors (22 cases).} Resource management issues are found more frequently
% 		in programming languages that do not manage memory automatically (e.g., C/C++
% 	    and Objective-C), i.e., where developers are responsible for
% 	    handling it instead. It is one of the main causes of DoS attacks.
% %
% 	\item \textbf{CWE-79: Improper Neutralization of Input During Web Page Generation/Cross-site Scripting (62 cases)} Lack of proper validation or escaping
% 	    allows attackers to submit untrusted data to web browsers through malicious
% 	    scripts that can hijack the user sessions or redirect the user to malicious
% 	    sites.
% %
% 	\item \textbf{CWE-22: Improper Limitation of a Pathname to a Restricted Directory/Path Traversal (37 cases).}  When developers forget to neutralize special elements within the pathnames to files/directories, attackers
% 	may leverage this flaw to resolve to a location outside of the restricted directory and gain access to documents
% 	or entire repositories.
% %
% 	\item \textbf{Miscellaneous (673 cases)} This pattern comprises
% 		several other security patches that do not have a CWE assigned and patterns that
% 		do not satisfy the Wilcoxon test size requirement of more than $20$ patches.
% \end{itemize}
%
\subsection{Security Patches vs. Regular Changes}
%
Previous studies attempted to measure the impact of regular changes 
on open-source software maintainability~\cite{HEGEDUS2018313}. 
However, there is no previous work focused on comparing the impact 
of security patches with regular changes on maintainability---only
with bug-fixes~\cite{10.1145/3133956.3134072}.
We analyze the maintainability of regular changes---changes not 
related to security patches---and, use them as a baseline.
The baseline dataset is generated from the security commits dataset, i.e.,
for each security commit in the dataset we collect a random regular
change from the same project. We created two different baselines: 
\textit{random-baseline},
considering random changes and all their characteristics; and,
\textit{size-baseline}, considering also random changes
but with the same size as security patches.

\subsubsection{Size-Baseline}

As for the security patches, for each regular change or refactoring, we also 
need the commit performing the regular change (\emph{sha-reg}) and version of 
the software before the change (\emph{sha-reg-p}). For each security patch, a 
\emph{diff} with an approximate 
size was fetched to a regular change. Due to the complexity of some patches, it was not possible 
to find patches with the exact number of added and deleted lines. 
Our search uses the security patch \emph{diff} to look for a 
\emph{diff} of the same size. The algorithm integrates the following 
steps:

@@

The pair of the random commit and its parent is accepted if the 
\emph{diff} size fits in the range size. This range widens every 
$10$ attempts to search for a change with the same size. We 
originate the regular changes from the security commits to ensure 
that differences in maintainability are not a consequence of 
characteristics of different projects.

\subsubsection{Random-Baseline}

As for the security patches, for each regular change or refactoring, we also 
need the commit performing the regular change (\emph{sha-reg}) and version of 
the software before the change (\emph{sha-reg-p}). For each security patch, 
a random commit is selected from the the project (\emph{sha-reg}).

@@

SIG---the company behind BCH--has been helping business and technology 
leaders drive their organizational 
objectives by fundamentally improving the health and security of 
their software applications for more than $20$ years. The
inner-working of their SIG-MM model---which is the one behind BCH---was 
scientifically proven~\cite{4335232,5609747,6113040,baggen2012}.
BCH not only considers security on their maintainability model~\cite{6616351}
but also integrates a total of $10$ different code metrics; and, 
code metrics were empirically validated in previous 
work~\cite{Bijlsma:2012:FIR:2317098.2317124,8530041,8919169,8785997}.

As BCH there are other tools that do the same kind of code analysis.
Two examples are Kiuwan and SonarCloud. Kiuwan is not free, their 
inner-workings do not seem to be published anywhere and there is
no clear list of metrics available---we inspected their datasheet.
SonarCloud is free but uses only $4$ of the guidelines (duplication,
size, complexity and code smells) that BCH considers. It also scores 
maintainability. In the following table, we
present a brief comparison between BCH and SonarCloud:

\Sofia{Currently building a script to retrieve the SonarCloud results}

\subsection{Maintainability Analysis}\label{sec:main_analysis}

In this research, we follow a very similar methodology to the one 
presented in previous work on the maintainability of energy-oriented 
fixes~\cite{8919169}. The inner-workings of BCH were proposed
originally in 2007~\cite{4335232} and suffer refinements later~\cite{5609747,6113040,baggen2012}. 
As said before, the web-based source code 
analysis service \emph{Better Code Hub} (BCH) is used to collect the 
maintainability reports of the patches of each project. 
Table~\ref{tab:guidelines} presents the $10$ guidelines proposed
by BCH's authors for delivering software that is not difficult to
maintain~\cite{Visser:2016:OREILLY} and, maps each guideline to the 
metric calculated by BCH. Theses guidelines are calculated using the 
metrics presented in~\cite{criteria:2017} and are also briefly 
explained on Table~\ref{tab:guidelines}. 
During each guideline evaluation, the tool determines the compliance 
towards one guideline by establishing limits for the percentage of 
code allowed to be in each of the $4$ risk severity levels
(\emph{low risk}, \emph{medium risk}, \emph{high risk}, and 
\emph{very high risk}). If the project does not violate those 
thresholds, then the BCH considers that the code is compliant with 
a guideline. These thresholds are determined by BCH using their own
data/experience---using open-source and closed software systems. If 
a project is compliant with a guideline, it means that it is at 
least $65\%$ better than the software used by BCH to calculate the 
thresholds\footnote{Check the answer to \emph{How can I adjust the 
threshold for passing/not passing a guideline?} at
\url{https://bettercodehub.com/docs/faq} (Accessed on \today{})}.

\begin{table}[h]
\centering
\scriptsize
	\caption{Guidelines to produce maintainable code. \Sofia{Still need to add more references}}
\begin{tabular}{L{2cm}L{4.5cm}L{4cm}}

@@

For projects with large codebases, the results retrieved for 
\emph{Keep Your Codebase Small} were not viable because they 
retrieved values above the limit set by BCH ($20$ Person-years). The 
\emph{Automated Tests} guideline was also not considered since the 
tool does not contemplate two of the most important techniques to 
security testing: vulnerability scanning and penetration testing. 
Instead, it only contemplates unit testing. Due to BCH limitations, 
we detected $308$ data points that retrieved incorrect overall 
calculations. Those data points were disregarded from the study.

@@

Many studies have investigated the relationship between patches and
software quality. Previous work focused on object-oriented metrics has evaluated the
impact of patches and exhibited proof that quantifying effectively the
impact of patches on maintainability may help to choose the appropriate
patch type~\cite{1167822}. In contrast to this work, Hegedus et
al.~\cite{HEGEDUS2018313} did not select particular metrics to assess the effect
of patches. Instead, statistical tests were used to find the metrics that
have the potential to change significantly after patches.

Researchers
performed a large-scale empirical study to understand the characteristics of security patches
and their differences against security bug fixes~\cite{10.1145/3133956.3134072}.
The main findings were that security patches are smaller and less complex compared
to bug fixes and usually performed at function-level.

Studying the evolution 
of maintainability issues during the development of Android apps, Malavolta et al. ($2018$)~\cite{8530041}
discovered that maintainability decreases over time. Palomba et al.
($2018$)~\cite{Palomba:2018:DIM:3231288.3231337} exhibits proof that code smells
should be carefully monitored by programmers since there is a high correlation
between maintainability aspects and proneness to changes/faults. In 2019, Cruz et 
al.~\cite{8919169} proposed a formula to calculate maintainability 
based on the BCH's guidelines and measured the impact of energy-oriented fixes 
on software maintainability. Recent work,
proposed a new maintainability model to 
measure fine-grained code changes by adapting/extending the BCH model~\cite{8785997}.
The present work uses the same model (SIG-MM), considers more guidelines and focuses 
solely on evaluating the impact of security patches on software maintainability.
